---
output:
  pdf_document: default
  html_document: default
---
--
title: "PSTAT126 Group Project Step 4"
author: "Hanya Ansari, Carina Yuen, Daren Aguilera"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2023-12-10"
---

## Introduction:

Wine Quality Based on Physicochemical Tests from UCI Machine Learning Repository <https://www.kaggle.com/datasets/uciml/red-wine-quality-cortez-et-al-2009/code?datasetId=4458&searchQuery=R>

No Missing Attribute Values: 0

Number of Instances: red wine: 1599

Number of Variables: 12 total, 11 continuous, 1 discrete (fixed_acidity, volatile_acidity, citric_acid, residual_sugar, chlorides, free_sulfur_dioxide, total_sulfur_dioxide, density, pH, sulphates, alcohol and 1 integer output variable: quality score between 0 and 10)

```{r}
knitr::opts_chunk$set(echo = FALSE,
                      message = F,
                      warning  = F,
                      fig.width = 4,
                      fig.height = 3,
                      fig.align = 'center',
                      fig.pos = 'H',
                      warning = FALSE)

library(leaps)
library(tidyverse)
library(tidymodels)
library(modelr)
library(ggplot2)
library(GGally)
library(olsrr)
library(glmnet)
library(trafo)
library(rsample)

#import data
wine_data <- read.csv("C:/Users/Carina W Yuen/Downloads/archive/winequality-red.csv")
#data <- read.csv("C:/Users/seren/Downloads/winequality-red.csv",sep=";")  
# read_delim("winequality-red.csv", delim=";")

#initialize variables
quality_data <- wine_data$quality
x_f <- wine_data$fixed.acidity
x_v <- wine_data$volatile.acidity
x_c <- wine_data$citric.acid
x_r <- wine_data$residual.sugar
x_ch <- wine_data$chlorides
x_fs <- wine_data$free.sulfur.dioxide
x_ts <- wine_data$total.sulfur.dioxide
x_d <- wine_data$density
x_p <- wine_data$pH
x_s <- wine_data$sulphates
x_a <- wine_data$alcohol
```

##Best model from step 3


```{r}
set.seed(112121)
wine_data
wine_partition <- resample_partition(wine_data,p = c(test = 0.3,train = 0.7))


fit_model_3 <- lm(x_v ~., data=wine_partition$train)
wine_partition_final<-wine_partition$train # without discrete variable quality score
out <- regsubsets(volatile.acidity~ ., data=wine,
               
                  method = 'seqrep',
                  nbest = 10,
                  nvmax = 10)
summary(out)
library(leaps)
library(tibble)
library(dplyr)
```


## Shrinkage Methods: Lasso and Ridge Regression with continuous response variable 
From our previous Project Step, our best model was fit_normal3, a linear model that has the response variable as volatile acidity, and the predictors (exlcuding itself) as the whole model. 


# Ridge Regression
Ridge regression shrinks the coefficients towards zero. Upon using coefficients (best_model), we got the following estimates: 10.42 for the intercept, 0.987 for alcohol, -0.0025 for free sulfur dioxide, and -0.0043 for total sulfur dioxide. Using cv.glmnet, we found that the optimal lambda is 0.03105, as a dotted vertical line marks the value of Log(λ) that minimizes the MSE value.

```{r}
require(glmnet)
```

```{r}
#x = model.matrix(Salary~., Hitters)[,-1]
#y = Hitters$Salary
x = model.matrix(quality~., data)[,-1]
y = wine_data$quality
cv_model0 <- cv.glmnet(x,y,alpha=1)

better_lambda <- cv_model0$lambda.min;better_lambda

better_model <- glmnet(x,y,alpha = 1, lambda = better_lambda)
coef(better_model)
```

```{r}
model1 <- lm(quality ~ -1.025008640*volatile.acidity + -0.005994492*citric.acid +0.002477866*residual.sugar + -1.762098962*chlorides + 0.002860910*free.sulfur.dioxide  + -0.002816664*total.sulfur.dioxide + -0.401321138*pH + 0.831568363*sulphates + 0.285812804*alcohol, data)
```


```{r}
# set grid 
grid = 10^seq(10, -2, length = 100)
ridge_mod = glmnet(x, y, alpha = 0, lambda = grid)
pred <- predict(object = ridge_mod, newx = x)
```

```{r}
set.seed(1) #we set a random seed first so our results will be reproducible.
cv.out.ridge=cv.glmnet(x, y, alpha = 0)
plot(cv.out.ridge)
abline(v = log(cv.out.ridge$lambda.min), lwd=3, lty=2)
bestlam = cv.out.ridge$lambda.min
bestlam
out = glmnet(x,y,alpha=0)
#predict(out,type="coefficients",s=bestlam)[1:20,]
```



```{r}
fit_normal3 <- lm(volatile.acidity~ ., data)
x_drop_va <- subset(data, select=-volatile.acidity)
scaled_x <- scale(x_drop_va) # drop volatile acidity
cv_model <- cv.glmnet(scaled_x, y, alpha=1)
best_lambda <- cv_model$lambda.min;best_lambda

print(best_lambda)

best_model <-glmnet(scaled_x, y, alpha = 0, lambda = best_lambda)
coefficients(best_model)

coef_pos <- abs(coefficients(best_model)[2:12])
coef_mean <- mean(coef_pos)
coef_mean

# plot?
par(mar = c(7, 4, 2.2, 0.5));plot(cv_model, cex=0.8)

```
# Lasso Regression
We decided to fit the other predictors to our response variable, volatile acidity. Using Lasso Regression, we obtained the following estimates for the intercept, and coefficients of fixed acidity, and density, respectively from the best model : 10.4230, 0.8485, -1.0751). Fixed acidity and density were highlighted because they were relatively large in magnitude, compared to the mean predictor magnitude (excluding the intercept) is about 0.3245. The optimal lambda (to minimize test MSE) was calculated to be about 0.001139. The small eigenvalue indicates multicollinearity.
```{r}
require(glmnet)

fit_normal3 <- lm(volatile.acidity~ ., data)
x_drop_va <- subset(data, select=-volatile.acidity)
scaled_x <- scale(x_drop_va) # drop volatile acidity
cv_model <- cv.glmnet(scaled_x, y, alpha=1)
best_lambda <- cv_model$lambda.min;best_lambda

print(best_lambda)

best_model <-glmnet(scaled_x, y, alpha = 1, lambda = best_lambda)
coefficients(best_model)

coef_pos <- abs(coefficients(best_model)[2:12])
coef_mean <- mean(coef_pos)
coef_mean

# plot?
par(mar = c(7, 4, 2.2, 0.5));plot(cv_model, cex=0.8)


# calculate R squared of model on the training data

# predictions using fitted best model
# va_predicted <- predict(cv_model, s=best_lambda, newx)
```
**Weighted Least Squares**
```{r}
plot(fitted(fit_normal3), resid(fit_normal3), xlab='fitted vals', ylab='residuals')

abline(0,0)

wt <- 1 / lm(abs(fit_normal3$residuals) ~ fit_normal3$fitted.values)$fitted.values^2

```

 
